---
title: "Team 8 - Assignment 1"
output:
  pdf_document: default
  html_document: default
date: "2024-02-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal - Segment the market & develop a targeting strategy using RFM analysis for Tuscan Lifestyles

```{r}
# Importing Packages
library(data.table)
library(psych)
library(ggplot2)
library(Hmisc)
```

```{r}
# Reading the Data
tuscan_data <- fread("TuscanDataForRFMAnalysis.csv",select = c(2,3,4,5,6,7))
cat("Data Columns: ",names(tuscan_data),"\n")
cat("No. of Sample Customers: ", nrow(tuscan_data),"\n")
```

## Question 1

What percentage of customers bought from the test catalog? Show the summary statistics (mean and standard deviation) of how much each buyer spent?

```{r}
pct_bought <- (nrow(tuscan_data[tuscan_data$buyerdummy == 1,])/nrow(tuscan_data))
cat("Percentage of Customers Bought From Test Catalogue: ",round(pct_bought*100,2),"%\n")
cat("Number of Customers Bought From Test Catalogue: ",nrow(tuscan_data[tuscan_data$buyerdummy == 1,]))
```

```{r}
buyer_data = tuscan_data[tuscan_data$buyerdummy == 1,]

mean_bought <- mean(buyer_data$dollars)
std_bought <- sd(buyer_data$dollars)
median_bought <- median(buyer_data$dollars)

cat("\nData Description\n")
describeBy(tuscan_data[,c("numords","totdol","last","dollars")],tuscan_data$buyer)

cat("Mean of Dollars Spent by Buyers: ", round(mean_bought,2),"\n")
cat("Standard Deviation of Dollars Spent by Buyers: ",round(std_bought,2),"\n")
cat("Median of Dollars Spent by Buyers: ",round(median_bought,2),"\n")

cat("\nCorrelation Plot\n")
rcorr(as.matrix(tuscan_data[,c('numords','totdol','last', 'dollars')]))
```

## Question 2

Create the quintiles for Recency (R), Frequency (F) and Monetary Value (M). Show 5 observations with their R, F, and M values and quintiles.

```{r}
tuscan_data$rec_iq <-.bincode(tuscan_data$last, quantile(tuscan_data$last, probs = seq(0, 1, 0.2)), right = TRUE,
include.lowest = TRUE)
describeBy(tuscan_data$last, tuscan_data$rec_iq)

tuscan_data$feq_iq <-.bincode(tuscan_data$numords, quantile(tuscan_data$numords, probs = seq(0, 1, 0.2)), right = TRUE,
include.lowest = TRUE)
describeBy(tuscan_data$numords, tuscan_data$feq_iq)

tuscan_data$mon_iq <-.bincode(tuscan_data$totdol, quantile(tuscan_data$totdol, probs = seq(0, 1, 0.2)), right = TRUE,
include.lowest = TRUE)
describeBy(tuscan_data$totdol, tuscan_data$mon_iq)
```


```{r}
cat("\nDisplayed 5 Random Rows\n")
tuscan_data[sample(nrow(tuscan_data),5),]
```

## Question 3: Display the averages of R, F, and M for each quintile in a bar chart.

```{r}
ggplot(data = tuscan_data, aes(x = rec_iq, y = last)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "skyblue") +
  geom_text(stat = "summary", aes(label = round(..y.., 2)), vjust = -0.5, color = "black") +
  labs(x = "Recency Quintile", y = "Days Since Previous Purchase (Recency)", title = "Recency Quintile Averages") +
  theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggplot(data = tuscan_data, aes(x = feq_iq, y = numords)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "pink") +
  geom_text(stat = "summary", aes(label = round(..y.., 2)), vjust = -0.5, color = "black") +
  labs(x = "Frequency Quintile", y = "Total Number of Past Purchases (Frequency)", title = "Frequency Quintile Averages") +
  theme_minimal() +
theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


ggplot(data = tuscan_data, aes(x = mon_iq, y = totdol)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "lightgreen") +
  geom_text(stat = "summary", aes(label = paste0("$ ",round((..y..), 2))), vjust = -0.5, color = "black") +
  labs(x = "Monetary Quintile", y = "Total Dollar Value of Past Purchases (Monetary)", title = "Monetary Quintile Averages") +
  theme_minimal() +
theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

## Question 4: Draw bar charts showing how response probability varies by R, F, and M quintiles. Comment briefly on what you learn from these bar charts.

```{r}
ggplot(data = tuscan_data, aes(x = rec_iq, y = buyerdummy)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "skyblue") +
  geom_text(stat = "summary", aes(label = paste0(round((..y..)*100, 2), "%")), vjust = -0.5, color = "black") +
  labs(x = "Recency Quintile", y = "Response Probability", title = "Recency Quintile Response Probability Averages") +
  theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

ggplot(data = tuscan_data, aes(x = feq_iq, y = buyerdummy)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "pink") +
  geom_text(stat = "summary", aes(label = paste0(round((..y..)*100, 2), "%")), vjust = -0.5, color = "black") +
  labs(x = "Frequency Quintile", y = "Response Probability", title = "Frequency Quintile Response Probability Averages") +
  theme_minimal() +
theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


ggplot(data = tuscan_data, aes(x = mon_iq, y = buyerdummy)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "lightgreen") +
  geom_text(stat = "summary", aes(label = paste0(round((..y..)*100, 2), "%")), vjust = -0.5, color = "black") +
  labs(x = "Monetary Quintile", y = "Response Probability", title = "Monetary Quintile Response Probability Averages") +
  theme_minimal() +
theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```
```{r}
tuscan_data$rec_iq <- 6-tuscan_data$rec_iq

ggplot(data = tuscan_data, aes(x = rec_iq, y = buyerdummy)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "skyblue") +
  geom_text(stat = "summary", aes(label = paste0(round((..y..)*100, 2), "%")), vjust = -0.5, color = "black") +
  labs(x = "Recency Quintile", y = "Response Probability", title = "Recency Quintile Averages") +
  theme_minimal() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
```{r}
tuscan_data$rfmindex_iq <- 100 * tuscan_data$rec_iq + 10 *tuscan_data$feq_iq + tuscan_data$mon_iq
tuscan_data$RFM_response <- ave(tuscan_data$buyerdummy, tuscan_data$rfmindex_iq)

ggplot(data = tuscan_data, aes(x = as.factor(rfmindex_iq), y = buyerdummy)) +
  geom_bar(stat = "summary", fun.y = "mean", fill = "red") +
  labs(title = "Respone Probability by RFM Index", x = "RFM Index", y = "Response Probability") +
  scale_x_discrete() +
  geom_hline(yintercept = 0.0246, color = "black", size = 1.5) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## Question 5: For this and the remaining questions, use the following numbers:

-   1,834,469 remaining customers (excluding the 5% used in the test)
-   Cost of goods and services = 50% of revenue
-   Cost of mailing an offer = $1
-   Average revenue per customer from those who order = (you calculated this in Q1)

##Suppose that Tuscan Lifestyles sends a mailing to all 1,834,469 remaining customers. Calculate the profitability of this mass-marketing strategy.


```{r}
# Assumption - Sample tested is representative of general population and can be used to extrapolate findings
# Assumption - Conservative approach - rounding down expected customers as integers 
no_customers<- 1834469
response_rate <- pct_bought
revenue_per_customer <- mean_bought

cost_mailing <- 1
cogs_per_customer<- revenue_per_customer*0.5
```

1.  How many buyers should you expect?

```{r}
exp_customer<- floor(no_customers*response_rate)
cat("Mass Marketing Expected No. of Customers: ",exp_customer,"\n")
```

2.  Net profit?

```{r}
total_cost<- (no_customers*cost_mailing) + (exp_customer*cogs_per_customer)
total_revenue<- exp_customer*revenue_per_customer
net_profit <- total_revenue - total_cost
cat("Mass Marketing Expected Net Profit: $",net_profit,"\n")
```

3.  Return on marketing expenditure (ROI?)

```{r}
mkt_roi<- net_profit/(no_customers*cost_mailing) 
cat("Mass Marketing Expected ROI - Marketing Expenditure: ",round(mkt_roi*100,2),"%\n")
```

## Question 6: Calculate the break-even response rate.

```{r}
break_even<- cost_mailing/(revenue_per_customer-cogs_per_customer)
cat("Mass Marketing Required Break Even Response Rate: ",round(break_even*100,2),"%\n")
```

## Question 7: Using an independent-quintile RFM approach and the break-even response rate, calculate:

1.  Optimum fraction of the market to target
```{r}
tuscan_data$target_iq[tuscan_data$RFM_response > break_even] <- 1
tuscan_data$target_iq[tuscan_data$RFM_response <= break_even] <- 0

target_fraction <- nrow(tuscan_data[tuscan_data$target_iq == 1])/nrow(tuscan_data)
cat("RFM Segmentation Optimum Fraction Target Market: ",round(target_fraction*100,2),"%\n")
```

```{r}
describeBy(tuscan_data$buyerdummy,tuscan_data$target_iq)
```
2.  The expected response rate
```{r}
rfm_response_rate <- mean(tuscan_data$RFM_response[tuscan_data$target_iq==1])
cat("RFM Segmentation Expected Response Rate: ",round(rfm_response_rate*100,2),"%\n")
```
3.  Net profit
```{r}
no_customers_rfm <- no_customers*target_fraction
exp_customer_rfm<- floor(no_customers_rfm*rfm_response_rate)

total_cost_rfm <- (no_customers_rfm*cost_mailing) + (exp_customer_rfm*cogs_per_customer)
total_revenue_rfm <- exp_customer_rfm*revenue_per_customer

net_profit_rfm <- total_revenue_rfm - total_cost_rfm
cat("Mass Marketing Expected Net Profit: $",net_profit_rfm,"\n")
```
4.  ROI of marketing expenditure
```{r}
mkt_roi_rfm<- net_profit_rfm/(no_customers_rfm*cost_mailing) 
cat("RFM Segmentation Expected ROI - Marketing Expenditure: ",round(mkt_roi_rfm*100,2),"%\n")
```
5.  Briefly explain why the RFM approach yields more profit and ROI than the mass-marketing approach

Refer to report submitted.

```{r}
CustomersRFMSegments <- table(tuscan_data$rfmindex_iq)
BuyersRFMSegments <-tapply(tuscan_data$buyerdummy, tuscan_data$rfmindex_iq, sum)
combined_table <- cbind(CustomersRFMSegments, BuyersRFMSegments)

# Convert to data frame
combined_df <- as.data.frame(combined_table)

write.csv(combined_df, file = "RFM_Segments.csv", row.names = TRUE)
```

